Reanalysis Step 2: Multiple Imputation

### 2025-08-29 reanalysis-2-multiple_imputation.qmd
bugs!

df_mice <- df_for_imputation |>
  mutate(
    # Create a unique row identifier to preserve exact row matching
    row_id = row_number()
  ) |>
  dplyr::select(row_id, dplyr::all_of(imputation_vars)) |>
  mutate(
    # Convert person_dr to ordered factor for polr imputation method
    person_dr = factor(person_dr, levels = c(0, 1, 2), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Lock factor levels for categorical variables to ensure consistency
    cohort_id = factor(cohort_id),  # Ensure cohort_id is a factor
    gender_cat = factor(gender_cat),
    race_ethnic_cat = factor(race_ethnic_cat),
    insurance_cat = factor(insurance_cat)
  )
```

@@

# Temporal VA variables - use 2-level PMM for clustered data
# 2l.pmm from miceadds: Multilevel PMM using linear mixed models
# This properly accounts for correlation between eyes from same patient
custom_methods["baseline_VA_logMAR"] <- "2l.pmm"  # Multilevel PMM for eyes within patients
custom_methods["intermediate_VA_logMAR"] <- "2l.pmm"  # Multilevel PMM for eyes within patients
custom_methods["outcome_VA_logMAR"] <- "2l.pmm"  # Multilevel PMM for eyes within patients
custom_methods["cohort_id"] <- ""  # Never impute - data management indicator only

>>>>

df_mice <- df_for_imputation |>
  mutate(
    # Create a unique row identifier to preserve exact row matching
    row_id = row_number()
  ) |>
  dplyr::select(row_id, dplyr::all_of(imputation_vars)) |>
  mutate(
    # Convert person_dr to ordered factor for polr imputation method
    person_dr = factor(person_dr, levels = c(0, 1, 2), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Lock factor levels for categorical variables to ensure consistency
    cohort_id = factor(cohort_id),  # Ensure cohort_id is a factor
    gender_cat = factor(gender_cat),
    race_ethnic_cat = factor(race_ethnic_cat),
    insurance_cat = factor(insurance_cat)
  ) |>
  mutate(
    # Clean factor levels to remove spaces and special characters for 2l.pmm compatibility
    # Spaces in factor levels cause formula parsing errors in lmer
    # Replace spaces and special characters with underscores
    across(where(is.factor), ~ {
      if (is.factor(.x)) {
        factor(.x, labels = gsub("[^[:alnum:]_]", "_", levels(.x)))
      } else {
        .x
      }
    })
  )
```

@@

# Temporal VA variables - use 2-level PMM for clustered data
# 2l.pmm from miceadds: Multilevel PMM using linear mixed models
# This properly accounts for correlation between eyes from same patient
# NOTE: 2l.pmm requires clean variable names (no spaces) in factor levels
custom_methods["baseline_VA_logMAR"] <- "2l.pmm"  # Multilevel PMM for eyes within patients
custom_methods["intermediate_VA_logMAR"] <- "2l.pmm"  # Multilevel PMM for eyes within patients
custom_methods["outcome_VA_logMAR"] <- "2l.pmm"  # Multilevel PMM for eyes within patients
custom_methods["cohort_id"] <- ""  # Never impute - data management indicator only
